---
title: "Map of Server Logins"
author: "Brian High"
date: "12/03/2015"
output:
  html_document:
    keep_md: yes
  pdf_document:
    fig_caption: no
    fig_height: 6
    fig_width: 9
---

## Set up our environment

Set document rendering options.
 
```{r set_options, echo=TRUE, message=FALSE}
# Configure `knitr` options.
library(knitr)
opts_chunk$set(tidy=FALSE, cache=FALSE)
```

Install packages if needed.

```{r install_packages, echo=TRUE, message=FALSE}
# Install packages (if necessary)
for (pkg in c("dplyr", "pander", "ggmap")) {
    if (! suppressWarnings(require(pkg, character.only=TRUE)) ) {
        install.packages(pkg, repos="http://cran.fhcrc.org", dependencies=TRUE)
        if (! suppressWarnings(require(pkg, character.only=TRUE)) ) {
            stop(paste0(c("Can't load package: ", pkg, "!"), collapse = ""))
        }
    }
}
```

## Read the data

```{r get_data}
cities <- read.csv("cities.csv")
```

## Take a look at the data

See which cities in which countries we are trying to map. List the top cities.

```{r top_ten_cities, results='asis'}
# Top-ten cities by login count
cities[,1:4] %>% head(10) %>% pandoc.table(style='rmarkdown')
```

Most logins are from Seattle, but what is the percentage of Seattle logins?

```{r seattle_logins}
# Percent of logins from Seattle
percent_seattle <- round(
    sum(cities[cities$city == "Seattle" & cities$region == "WA", "logins"]) / 
    sum(cities[, "logins"]) * 100, 1)
percent_seattle
```

List the top countries.

```{r top_ten_countries, results='asis'}
# Top-ten countries by login count
cities %>% group_by(country) %>% 
    summarize(logins=sum(logins)) %>% arrange(desc(logins)) -> countries
countries %>% head(10) %>% pandoc.table(style='rmarkdown')
```

Find percent of logins from the US.

```{r us_logins}
# Percent of logins from US
percent_us <- round(
    sum(cities[cities$country == "US", "logins"]) / 
    sum(cities[, "logins"]) * 100, 1)
percent_us
```

So, `r percent_seattle` % of the logins are from Seattle and `r percent_us` %
are from the US. 

## Make the map

We can center the map on the center of the US (Kansas) and zoom so that we 
pick up some of the locations in the rest of North America and Central America.

```{r make_google_map, warning=FALSE, message=FALSE}
# Center the map on Kansas
kansas <- geocode("Kansas")
google_basemap <- get_map(location = c(lon = kansas$lon, lat = kansas$lat), 
                          zoom = 4, scale=2)

ggmap(google_basemap) + geom_point(data = cities, 
                                   aes(x = longitude, y = latitude,
                                fill = "red", alpha = 0.8), 
                                size = (cities$logins)^.5, shape = 21) + 
    guides(fill=FALSE, alpha=FALSE, size=FALSE)
```

Map and geocoding sources: 

- _ggmap_. version 2.5.2. David Kahle and Hadley Wickham. [Web 3 Dec. 2015](https://github.com/dkahle/ggmap).
- _Google Maps_. Google, 3 Dec. 2015. [Web 3 Dec. 2015](http://maps.googleapis.com/maps/api/staticmap?center=37.697948,-97.314835&zoom=4&size=640x640&scale=2&maptype=terrain&language=en-EN&sensor=false).  
